---
title: "LPS Sensitivity"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---


The purpose of this workflow is to start exploring the LPS sensitivity
data and to run cape on it to see if there any informative interactions.



```{r setup}
rm(list = ls())

library(here)

all.fun <- list.files(here("Code"), full.names = TRUE, pattern = ".R")
for(i in 1:length(all.fun)){source(all.fun[i])}

all.packages <- c("pheatmap", "qtl2", "cluster")
load_libraries(all.packages)

load_latest_cape(here("../../../git_repositories/cape"))
```


We need to take the non-numeric phenotypes out of the phenotype
file and put them into covariates. I also added 'NA' to the na.strings
argument in the yaml file.

```{r num_pheno}
#if this is the first time through, create a new data directory

mod.data.dir <- here("Data", "qtl2_files_mod")

if(!file.exists(mod.data.dir)){
    #duplicate all files
    file.copy(here("Data", "qtl2_files"), mod.data.dir, recursive = TRUE)

    orig.pheno <- read.csv(here("Data", "qtl2_files", "lps_pheno.csv"))
    orig.covar <- read.csv(here("Data", "qtl2_files", "lps_covar.csv"))

    nonnum.pheno.col <- c("strain", "category")
    num.pheno.col <- setdiff(colnames(orig.pheno), nonnum.pheno.col)
    nonnum.pheno <- cbind(orig.covar, orig.pheno[,nonnum.pheno.col])

    num.pheno <- orig.pheno[,num.pheno.col]

    write.table(num.pheno, here("Data", "qtl2_files_mod", "qtl2_files", "lps_pheno.csv"),
        sep = ",", quote = FALSE, row.names = FALSE)
    write.table(nonnum.pheno, here("Data", "qtl2_files_mod", "qtl2_files", "lps_covar.csv"), 
        sep = ",", quote = FALSE, row.names = FALSE)
}
```

```{r load_data}
#add 'NA' to the na.strings arguments in the yaml file to 
#suppress the warning about "NAs introduced by coercion"
lps <- read_cross2(here("Data", "qtl2_files_mod", "qtl2_files", "lps.yaml"))
lps_cape <- qtl2_to_cape(lps)

data_obj <- lps_cape$data_obj
geno_obj <- lps_cape$geno_obj
```

## Body Temperature

Body temperature was measured hourly for seven hours post
LPS injection. The following plot shows a heatmap of the 
temperatures after the injection. There are several patterns.
Some temperatures stay fairly steady. Others dip a bit after
a few hours, and some really drop in the last couple hours
of the experiment. In some animals the temperatures go up
initially, and others hold more steady before dropping.

```{r explore_pheno, fig.width = 5, fig.height = 7}
temp.idx <- grep("X", colnames(data_obj$pheno))
temp.label <- gsub("X", "", colnames(data_obj$pheno)[temp.idx])
temp <- data_obj$pheno[,temp.idx]

pheatmap(temp, cluster_cols = FALSE)
```


```{r k}
k = 7
```

We can cluster the tamperature profiles hierarchically
to see the different patterns. The following heatmap
shows the cluster assignments for `r k` clusters.

```{r cluster_temp, fig.width = 5, fig.height = 7}

temp_clust <- hclust(dist(temp))
cl <- cutree(temp_clust, k = k)

cl_annot <- data.frame("cluster" = as.factor(cl))
rownames(cl_annot) <- rownames(temp)
pheatmap(temp, annotation_row = cl_annot, cluster_cols = FALSE)
```

If we plot the profiles out for each cluster independently,
we can better see the different profiles.

```{r cluster_sep, fig.width = 9, fig.height = 7}

layout.mat <- get.layout.mat(k)
layout(layout.mat)
for(cl.k in 1:k){
    plot.new()
    plot.window(xlim = c(1,ncol(temp)), ylim = c(min(temp),  max(temp)))
    cl.idx <- which(cl == cl.k)
    axis(1);mtext("Hrs", side = 1, line = 2.5)
    axis(2);mtext("Temp (C)", side = 2, line = 2.5)
    mtext(paste("Cluster", cl.k), side = 3)
    for(i in 1:length(cl.idx)){
        points(1:ncol(temp), temp[cl.idx[i],], col = cl[cl.idx[i]], type = "b", pch = 16)
    }
}

```

If we plot the first two principal components of the temperature
matrix against each other, we see that the clusters separate along
the first principal component, which accounts for 84% of the variation
in temperature. This suggests that the first prinicipal component
might be a good trait for mapping.

```{r temp_decomp, fig.width = 5, fig.height = 5}
temp.decomp <- plot.decomp(temp, cols = cl)
temp.pc <- temp.decomp$u[,1,drop=FALSE]
colnames(temp.pc) <- "Temp_PC1"
rownames(temp.pc) <- rownames(temp)
data_obj$pheno <- cbind(data_obj$pheno, temp.pc)
#boxplot(temp.decomp$u[,1]~cl)
```

## Trait Correlations

The animals in the upper ranges of the first PC of the temperature 
matrix, also tend to have higher measurements on in the immune
matrix. 

```{r immune, fig.width = 8, fig.height = 8}
immune <- data_obj$pheno[,c("Il.1b", "Il.6", "TNF", "IFN.b")]
norm.immune <- apply(immune, 2, rankZ)

par(mfrow = c(2,2))
for(i in 1:ncol(norm.immune)){
    plot.with.model(temp.pc, immune[,i], xlab = "Temperature PC 1",
        ylab = colnames(norm.immune)[i],
        main = colnames(norm.immune)[i], report = "cor.test",
        col = cl)
}
```

These correlations extend the range of the values. If we 
rank Z normalize the values, the correlations are even 
stronger. There is a floor effect in each of the immune 
molecules, which is most pronounce in IFN beta. There are 
animals with a range of temperature profiles that did 
not have measurable IFN beta.


```{r norm_immune, fig.width = 8, fig.height = 8}
par(mfrow = c(2,2))
for(i in 1:ncol(norm.immune)){
    plot.with.model(temp.pc, norm.immune[,i], xlab = "Temperature PC 1",
        ylab = paste("Normalized", colnames(norm.immune)[i]),
        main = paste("Normalized", colnames(norm.immune)[i]), report = "cor.test",
        col = cl)
}
```


```{r immune_scratch, eval = FALSE}
pheatmap(immune, scale = "column")
pairs(immune)
pairs(norm.immune)
pheatmap(cor(norm.immune), display_numbers = TRUE)
pheatmap(cor(immune), display_numbers = TRUE)

boxplot(immune[,4]~cl)
```


## Individual Trait Mapping {.tabset .tabset-fade .tabset-pills}

As an initial test, we mapped each of the normalized traits.

The first PC of the temperature matrix has a suggestive QTL
on proximal Chr 17. Maybe the MHC locus? IL6 and TNF also map 
to Chr 17, although the QTLs look more distal. Neither IL1 beta 
nor IFN beta map anywhere. 

```{r temp_scan, fig.width = 9, fig.height = 5, results = "asis"}
genoprobs <- calc_genoprob(lps)
full.pheno <- cbind(temp.pc, norm.immune)
temp.scan <- scan1(genoprobs, full.pheno)
temp.perm <- scan1perm(genoprobs, full.pheno, n_perm = 100)
alpha <- c(0.05, 0.01, 0.1)
sig.level <- summary(temp.perm, alpha = alpha)

for(ph in 1:ncol(full.pheno)){
    cat("###", colnames(full.pheno)[ph], "\n")
   plot(temp.scan, lodcol = ph, map = lps$pmap, ylim = c(0, max(c(max(temp.scan), max(sig.level)*1.05))),
        main = colnames(full.pheno)[ph])
    plot.dim <- par("usr")
    plot.width <- plot.dim[2] - plot.dim[1]
    text.x <- plot.dim[2]-(plot.width*0.007) 
    line.stop <- plot.dim[2]-(plot.width*0.04) 
    segments(x0 = 0, x1 = line.stop, y0 = sig.level[,ph], lty = 2)
    text(x = text.x, y = sig.level[,ph], labels = alpha, adj = 1)
    cat("\n\n")
}
```

## CAPE

Now we run cape with the combined temperature PC1 and immune
molecule matrix. Please see cape_results.html for results from
cape.

```{r cape}
results_dir <- here("Results", "cape_2ET")
final_obj <- run_cape(data_obj, geno_obj, 
    param_file = file.path(results_dir, "0_lps.parameters_0.yml"),
    results_path = results_dir)
```