---
title: "LPS Sensitivity"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---


The purpose of this workflow is to start exploring the LPS sensitivity
data and to run cape on it to see if there any informative interactions.



```{r setup}
rm(list = ls())

library(here)

all.fun <- list.files(here("Code"), full.names = TRUE, pattern = ".R")
for(i in 1:length(all.fun)){source(all.fun[i])}

all.packages <- c("pheatmap", "qtl2", "cluster")
load_libraries(all.packages)

load_latest_cape("../../../git_repositories/cape")
```


We need to take the non-numeric phenotypes out of the phenotype
file and put them into covariates. I also added 'NA' to the na.strings
argument in the yaml file.

```{r num_pheno}
#duplicate all files
file.copy(here("Data", "qtl2_files"), here("Data", "qtl2_files_mod"), recursive = TRUE)

orig.pheno <- read.csv(here("Data", "qtl2_files", "lps_pheno.csv"))
orig.covar <- read.csv(here("Data", "qtl2_files", "lps_covar.csv"))

nonnum.pheno.col <- c("strain", "category")
num.pheno.col <- setdiff(colnames(orig.pheno), nonnum.pheno.col)
nonnum.pheno <- cbind(orig.covar, orig.pheno[,nonnum.pheno.col])

num.pheno <- orig.pheno[,num.pheno.col]

write.table(num.pheno, here("Data", "qtl2_files_mod", "qtl2_files", "lps_pheno.csv"),
    sep = ",", quote = FALSE, row.names = FALSE)
write.table(nonnum.pheno, here("Data", "qtl2_files_mod", "qtl2_files", "lps_covar.csv"), 
    sep = ",", quote = FALSE, row.names = FALSE)

```

```{r load_data}
#add 'NA' to the na.strings arguments in the yaml file to 
#suppress the warning about "NAs introduced by coercion"
lps <- read_cross2(here("Data", "qtl2_files_mod", "qtl2_files", "lps.yaml"))
lps_cape <- qtl2_to_cape(lps)

data_obj <- lps_cape$data_obj
geno_obj <- lps_cape$geno_obj
```

Body temperature was measured hourly for seven hours post
LPS injection. The following plot shows a heatmap of the 
temperatures after the injection. There are several patterns.
Some temperatures stay fairly steady. Others dip a bit after
a few hours, and some really drop in the last couple hours
of the experiment. In some animals the temperatures go up
initially, and others hold more steady before dropping.

```{r explore_pheno, fig.width = 5, fig.height = 7}
temp.idx <- grep("X", colnames(data_obj$pheno))
temp.label <- gsub("X", "", colnames(data_obj$pheno)[temp.idx])
temp <- data_obj$pheno[,temp.idx]

pheatmap(temp, cluster_cols = FALSE)
```


```{r k}
k = 7
```

We can cluster the tamperature profiles hierarchically
to see the different patterns. The following heatmap
shows the cluster assignments for `r k` clusters.

```{r cluster_temp, fig.width = 5, fig.height = 7}

temp_clust <- hclust(dist(temp))
cl <- cutree(temp_clust, k = k)

cl_annot <- data.frame("cluster" = as.factor(cl))
rownames(cl_annot) <- rownames(temp)
pheatmap(temp, annotation_row = cl_annot, cluster_cols = FALSE)
```

If we plot the profiles out for each cluster independently,
we can better see the different profiles.

```{r cluster_sep, fig.width = 9, fig.height = 7}

layout.mat <- get.layout.mat(k)
layout(layout.mat)
for(cl.k in 1:k){
    plot.new()
    plot.window(xlim = c(1,ncol(temp)), ylim = c(min(temp),  max(temp)))
    cl.idx <- which(cl == cl.k)
    axis(1);mtext("Hrs", side = 1, line = 2.5)
    axis(2);mtext("Temp (C)", side = 2, line = 2.5)
    mtext(paste("Cluster", cl.k), side = 3)
    for(i in 1:length(cl.idx)){
        points(1:ncol(temp), temp[cl.idx[i],], col = cl[cl.idx[i]], type = "b", pch = 16)
    }
}

```

If we plot the first two principal components of the temperature
matrix against each other, we see that the clusters separate along
the first principal component, which accounts for 84% of the variation
in temperature. This suggests that the first prinicipal component
might be a good trait for mapping with.

```{r temp_decomp, fig.width = 5, fig.height = 5}
temp.decomp <- plot.decomp(temp, cols = cl)
temp.pc <- temp.decomp$u[,1,drop=FALSE]
colnames(temp.pc) <- "Temp_PC1"
rownames(temp.pc) <- rownames(temp)
data_obj$pheno <- cbind(data_obj$pheno, temp.pc)
#boxplot(temp.decomp$u[,1]~cl)
```

The animals in the upper ranges of the first PC of the temperature 
matrix, also tend to have higher measurements on in the immune
matrix. 

```{r immune, fig.width = 8, fig.height = 8}
immune <- data_obj$pheno[,c("Il.1b", "Il.6", "TNF", "IFN.b")]
norm.immune <- apply(immune, 2, rnorm)

par(mfrow = c(2,2))
for(i in 1:ncol(norm.immune)){
    plot.with.model(temp.pc, immune[,i], xlab = "Temperature PC 1",
        ylab = colnames(norm.immune)[i],
        main = colnames(norm.immune)[i], report = "cor.test",
        col = cl)
}
```

However, there is no correlation if we normalize the immune
measurements.

```{r norm_immune, fig.width = 8, fig.height = 8}
par(mfrow = c(2,2))
for(i in 1:ncol(norm.immune)){
    plot.with.model(temp.pc, norm.immune[,i], xlab = "Temperature PC 1",
        ylab = paste("Normalized", colnames(norm.immune)[i]),
        main = paste("Normalized", colnames(norm.immune)[i]), report = "cor.test",
        col = cl)
}
```

This suggests that there are a few outlier mice that have a 
huge immune response that also corresponds with a temperature
drop, but that the smaller variations in the immune molecules
do not correspond to the temperature response in general. 
This doesn't really bode well for pulling out interactions 
with CAPE, but we can still try.

```{r immune_scratch, eval = FALSE}
pheatmap(immune, scale = "column")
pairs(immune)
pairs(norm.immune)
pheatmap(cor(norm.immune), display_numbers = TRUE)
pheatmap(cor(immune), display_numbers = TRUE)

boxplot(immune[,4]~cl)
```

The first PC of the temperature matrix doesn't map anywhere
significantly, although the Chr 17 locus does look pretty 
good. Probably MHC?

```{r temp_scan, fig.width = 9, fig.height = 5}
genoprobs <- calc_genoprob(lps)
temp.scan <- scan1(genoprobs, temp.pc)
temp.perm <- scan1perm(genoprobs, temp.pc, n_perm = 100)
sig.level <- summary(temp.perm, alpha = c(0.05, 0.01, 0.1))

plot(temp.scan, map = lps$pmap, ylim = c(0, max(c(max(temp.scan), max(sig.level)*1.05))),
    main = "Temp PC1")
abline(h = sig.level[,1], lty = c(2,1))
```

Now we will run cape with the combined temperature PC1 and immune
molecule matrix.

```{r cape}
results_dir <- here("Results", "cape")
final_obj <- run_cape(data_obj, geno_obj, 
    param_file = file.path(results_dir, "0_lps.parameters_0.yml"),
    results_path = results_dir)
```